# Introduction {#intro}

The main purpose of this document is to show you how to analyze data from electrophysiological experiments using the *relectro* R package. It will describe the main functionalities of *relectro*. 

## Why R and *relectro*

There are several different alternative environments to do analysis of electropysiological data.
I decided to use R because it is distributed under GNU general public license. 
This means that it will always be available and that we don't depend on a company to give us the permission to use it.
In the context of scientific work, this means that anyone can replicate your analysis without having to buy software.

R also has other advantages. 
It has probably the best support for statistical analysis and it can also generate pretty nice figures.

It's main disadvantages are that it is pretty slow when compared to c codes, and some calculations are not easily done with the R language. 

This is where the *relectro* package kicks in. It provides a high-level interface for the analysis of electrophysiological data in R.
Under the hood, key calculations are implemented in c.
So you get the speed of c, but with a much simplified interface.

The other motivation for *relectro* is to promote the *re-use* of code. 
Therefore, we solve every bug only once, and the next user benefits from the work. 
There is no need (or time) for each lab member to reinvent the wheel.


## Scope

This document is not an introduction to R. 
If you have no experience with R, I would recommand doing a couple of tutorials online.
A very good book is [The Art of R Programming: A Tour of Statistical Software Design](https://www.amazon.de/Art-Programming-Statistical-Software-Design/dp/1593273843/ref=sr_1_1?ie=UTF8&qid=1481745150&sr=8-1&keywords=r+programming).

## Organization

This document is organized in chapters.
One topic is covered in each chapter and it assumes that you are familiar with the previous chapters.

Once you are familiar with the content, you can use this document as a quick reference.

## Let's get started

The first step is to launch RStudio. 
If you have not used RStudio before, you might want to read or watch a quick tutorial online.

The first thing you need to use *relectro* is to load it into your R session.
You do this by running the code in the gray box below.
This should be done each time you start a new R session. 
Otherwise, you will not have access to the functions provided by *relectro*
```{r intro}
library(relectro)
```

*relectro* is an object-oriented package. 
This means that as a user you will be creating objects and applying functions to these objects.
For example, you can create an object of the class RecSession to represent one recording session.
This might sound rather abstract but it will become clear in the next chapter.


You can have a look at all classes and functions that are part of *relectro*.
```{r listObjects,echo=TRUE,results='hide'}
ls("package:relectro")
```
I am not showing the results here to save space but you should see a list of more than 150 objects.
The names of the classes start with an upper case letter (e.g. RecSession) whereas the names of functions and methods start a lower case letter (e.g. loadRecSession).


The exsiting classes in *relectro* are
```{r listClasses,echo=FALSE}
ls("package:relectro")[grep("^[[:upper:]]",ls("package:relectro"))]
```

The next chapters cover these objects and show you how to manipulate them.

