# Modify relectro

relectro is far from being a very polished package. It might be that you have to modify it.
There are several reasons why you may want to make modifications to relectro:

1. Fix a bug
2. Add a new function
3. Add modify and existing function
4. Improve the documentation
5. Improve its running speed

Here we provide a step-by-step procedure to modify relectro on your computer and to get your changes in the main repository. This assumes that you are using RStudio but you don't have to.

## Fork relectro

If you want to make change to relectro, the best way is to fork the relectro git repository. 
The fork just create a copy of the relectro package in your github personal account. 

Then use `git clone yourForkRepoURL` to get a copy of the repo on your computer.
You are then free to experiment with your own code.

You can find some explanation of the fork process on [github](https://help.github.com/articles/fork-a-repo/).

## Install the devtools R package

Open RStudio and get the package
```{r, eval=FALSE}
install.packages("devtools")
```

## Load relectro as a project in RStudio

In RStudio go to `File -> Open Project`. Locate the file `relectro.Rproj`.

The RStudio environment will facilitate the way you can change relection, build and reload relectro and test your modification

## Make the changes

Modifying relectro usually involves a few cycles of this procedure:

1. Make modifications to the code
2. Build and reload (Build -> Build and Reload)
3. Test the code you wrote with some examples


## Update the documentation

If you make modifications to a function or add a function in relectro, you also need to update the documentation.
You have to use the roxygen2 package to do this.
```{r, eval=FALSE}
install.packages("roxygen2")
```

Add or modify the roxygen comments to your source file. You can find information on how to do this with the [roxygen vignette](https://cran.r-project.org/web/packages/roxygen2/vignettes/rd.html).

Create the Rd files for the documentation by running
```{r, eval=FALSE}
devtools::document()
```


## Add tests for your new function

If you add a function to the relectro package, you are strongly encouraged to test your function with formal tests.
To do so you will need the testthat package

```{r, eval=FALSE}
install.packages("testthat")
```

You can find more information about writing tests [here](http://r-pkgs.had.co.nz/tests.html).
All the tests are in the `tests/testthat` folder of the relectro repository. 


## Check that you haven't broken relectro

There is a series of checks that you can run to pick up frequent mistakes. 
```{r, eval=FALSE}
devtools::check(document = FALSE)
```
You can also run the same tests in RStudio with `Build -> Check Package`
This should exit without warnings or errors.


## Commit your changes to the forked repository (your fork)

Once you are happy with the modifications you made. Commit the changes using git.
Below is one way to do this.

```{r, engine = 'bash', eval = FALSE}
cd /home/kevin/repo/relectro
git commit -a -m "message explaining the modification done"
```


## Create a pull request

If you want to have your changes incorporated into the main relectro package, you just need to create a pull request on the main relectro repository.
You can get more information [here](https://help.github.com/articles/about-pull-requests/).

Once your changes are approved, they will be available to all users.

