# Deploy the relectro_doc book on an apache2 server

This document shows you how to deploy the book on a local apache2 server.

## Install Apache, MySQL, PHP

Choose a computer to host the book.
Log in this computer and install the 3 components.
It is assumed that you are using a recent ubuntu computer.

```{r, engine = 'bash', eval = FALSE}
sudo apt-get update
sudo apt-get upgrade
sudo apt-get install lamp-server^ phpmyadmin
```

Mind the ^ at the end.

## Test Apache server
Test that the Apache server works by opening a browser and go to the ip address of the computer you choose.
You should get a page with the message `It works!`.

## Test that php works

Create a file `/var/www/html/test.php` with this content.
```{r, engine= 'bash', eval = FALSE}
<? php_info(); ?>
```
The apache server should have the permission to read the file.
You should see a page describing your php engine.

## Allow mysql connection from php

Edit the file `/etc/php/7.0/apache2/php.ini`
Uncomment `extension=php_mysqli.dll`

## Create a directory for sensitive information

```{r, engine='bash',eval=FALSE}
sudo mkdir /var/www/html/includes
sudo chgrp data /var/www/html/includes
sudo chmod g+wrx /var/www/html/includes
```

## Create an empty mysql database

Replace `xxxx` and `yyyyyy` by your passwords.

```{r, engine='bash',eval=FALSE}
mysql -u root  -p
CREATE DATABASE dbtest;
grant all privileges on *.* to 'kevin'@'localhost' identified by 'xxxx'; 
GRANT SELECT, INSERT, UPDATE, DELETE, CREATE, DROP, INDEX, ALTER, CREATE TEMPORARY TABLES, LOCK TABLES ON dbtest.* TO 'webuser'@'localhost' IDENTIFIED BY 'yyyyyy';
quit;
```

## Create a table in the mysql database
```{r, engine='bash',eval=FALSE}
 mysql -u kevin -p dbtest
CREATE TABLE IF NOT EXISTS `users` (
  `userId` int(11) NOT NULL AUTO_INCREMENT,
  `userName` varchar(30) NOT NULL,
  `userEmail` varchar(60) NOT NULL,
  `userPass` varchar(255) NOT NULL,
  PRIMARY KEY (`userId`),
  UNIQUE KEY `userEmail` (`userEmail`)
) ENGINE=InnoDB DEFAULT CHARSET=utf8 AUTO_INCREMENT=1 ;
```

## Ideally we would need a php-mysql registering system but can't get it to work.

At the moment, I clone the repo on the server.
Run bookdown::render_book("index.Rmd")
sudo mv _book/* /var/www/html

This is very simple but there is no registration.

