# Deploy the relectro_doc book on an apache2 server

This document shows you how to deploy the book on a local apache2 server.

## Install Apache, MySQL, PHP

This is an overkill but in case we want to use a database in the future.
Choose a computer to host the book.
Log in this computer and install the 3 components.
It is assumed that you are using a recent ubuntu computer.

```{r, engine = 'bash', eval = FALSE}
sudo apt-get update
sudo apt-get upgrade
sudo apt-get install lamp-server^ phpmyadmin
```

Mind the ^ at the end.

## Test Apache server
Test that the Apache server works by opening a browser and go to the ip address of the computer you choose.
You should get a page with the message `It works!`.

## Test that php works

Create a file `/var/www/html/test.php` with this content.
```{r, engine= 'bash', eval = FALSE}
<? php_info(); ?>
```
The apache server should have the permission to read the file.
You should see a page describing your php engine.

## Set group permissiton of /var/www/html
```{r, engine= 'bash', eval = FALSE}
sudo chgrp www-data html
sudo chmod g+wrx html
sudo usermod -a -G www-data kevin
```

## Allow mysql connection from php

Edit the file `/etc/php/7.0/apache2/php.ini`
Uncomment `extension=php_mysqli.dll`

## Install pandoc version 1.17.2 or newer
```{r, engine = 'bash', eval = FALSE}
sudo apt-get install pandoc
pandoc --version
```

## Clone relectro and relectroDoc on server
```{r, engine= 'bash', eval =FALSE}
cd /home/kevin/repo/
git clone https://github.com/kevin-allen/relectroDoc.git
git clone https://github.com/kevin-allen/relectro.git
```

## Use crontab on the server to deploy

There is a shell script to deploy the relectroDoc html documentation: `/home/kevin/repo/relectroDoc/deploy/deploy_relectroDoc.sh`.
Install it in the crontab of user kevin (see comment in the shell script itself)

